---
title: "DATA 622 - Homework 2"
author: "OMER OZEREN"
output:
  html_document:
    highlight: tango
    theme: journal
    toc: yes
    toc_depth: 5
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '5'
---

```{r echo=FALSE, include=FALSE}
library(class)
library(pROC)
library(ROCR)
library(caret)
library(C50)
library(e1071)
library(naivebayes)
library(MASS)
library(mltools)
library(knitr)
```

## PART A

### Introduction
STEP#0: Pick any two classifiers of (SVM,Logistic,DecisionTree,NaiveBayes). Pick heart or ecoli dataset. Heart is simpler and ecoli compounds the problem as it is NOT a balanced dataset. From a grading perspective both carry the same weight.

STEP#1 For each classifier, Set a seed (43)

STEP#2 Do a 80/20 split and determine the Accuracy, AUC and as many metrics as returned by the Caret package (confusionMatrix) Call this the base_metric. Note down as best as you can development (engineering) cost as well as computing cost(elapsed time).
Start with the original dataset and set a seed (43). Then run a cross validation of 5 and 10 of the model on the training set. Determine the same set of metrics and compare the cv_metrics with the base_metric. Note down as best as you can development (engineering) cost as well as computing cost(elapsed time).
Start with the original dataset and set a seed (43) Then run a bootstrap of 200 resamples and compute the same set of metrics and for each of the two classifiers build a three column table for each experiment (base, bootstrap, cross-validated). Note down as best as you can development (engineering) cost as well as computing cost(elapsed time).

### Import Data 
```{r}
data <- read.csv("https://raw.githubusercontent.com/omerozeren/DATA624/master/HMW10/GroceryDataSet.csv")
```
### Split Data into Train(80%) and Test data(20%)  

```{r}
set.seed(998)
split_df <- createDataPartition(df$label, p = .80, list = FALSE)
df_train <- df[split_df,]
df_test <- df[-split_df,]
```